### 2019-05-06 Thesis/pmonads2.tex
Nightly build.

\begin{definition}
  Given a set $X$ and a \Mellies morphism $\sigma \from 1 \to (X \to \bC)$, we say that $\sigma$ \emph{accepts} a sequence $u\in X^*$ if $\sigma;\xi_u\ne\bot$.
  We write $\Acc(\sigma)$ for the set of all sequences accepted by $\sigma$.

  More generally, if $\sigma\from 1 \to (X \to A)$ is a \Mellies morphism and $\tau\from 1 \to A$ is an ordinary morphism in $\G$, we write
  \[
    \Acc_\tau(\sigma) = \{u\in X^*\suchthat \sigma;\xi_u = \tau\}\,.
    \]
  When $A=\bC$, we have $\Acc(\sigma) = \Acc_{\deno{\skipp}}(\sigma)$.
  \label{DefAcc}
\end{definition}

@@

\begin{proposition}
  Let $\sigma\from 1 \to (j(p) \to A)$, $\sigma'\from 1 \to (j(q) \to A)$ be two representatives of the same morphism $1 \to A$ in $\G/\oppcat\X$.  
  Let $\tau\from 1 \to A$ be an ordinary morphism in $\G$.
  Then $(p,\Acc_\tau(\sigma))$ and $(q,\Acc_\tau(\sigma'))$ are equivalent.
  In particular, if $A=\bC$, then $(p,\Acc(\sigma))$ and $(q,\Acc(\sigma')$ are equivalent.
  \label{PropAccEquivalent}
\end{proposition}
\begin{proof}
  Since the relation on pairs $(p,\U)$ is an equivalence relation, it suffices to assume that $\sigma$ and $\sigma'$ are related by the relation that generates the equivalence relation on \Mellies morphisms; i.e., that there is a morphism $f\from p \to q$ such that $\sigma = \sigma';(j(f)\to \bC)$.

  Let $u\in j(p)^*$.  
  Then we have
  \begin{IEEEeqnarray*}{rCl}
    u\in\Acc_\tau(\sigma) & \Leftrightarrow & \sigma;\xi_u=\tau \\
    & \Leftrightarrow & \sigma';(j(f)\to A);\xi_u=\tau \\
    & \Leftrightarrow & \sigma';\xi_{j(f)_*u}=\tau \\
    & \Leftrightarrow & j(f)_*u \in \Acc_\tau(\sigma')\,.
  \end{IEEEeqnarray*}
  Therefore, $(p,\Acc_\tau(\sigma))$ and $(q,\Acc_\tau(\sigma'))$ are equivalent.
\end{proof}

@@

@@

  Let $s\in\sigma\semicom_{\text{Mell}}\tau$.  
  Then there exist four sequence $s_X,s_Y,\s_X,\s_Y$ such that
  \begin{itemize}
    \item $s$ is an interleaving of the sequences $s_X$ and $s_Y$;
    \item $\s_X\in\sigma$ and $\s_Y\in\tau$;
    \item $s_X\vert_{A,X} = \s_X\vert_{A,X}$; and
    \item $s_Y\vert_{C,Y} = \s_Y\vert_{C,Y}$.
  \end{itemize}
  \label{LemMelliesDecomposition}
\end{lemma}
\begin{proof}
  Recall that $\sigma\semicom_{\text{Mell}}\tau$ is given by the following composite in $\G$, where $\mu$ is the composite $\langle\ul{\pr_X},\ul{\pr_Y}\rangle$.
  \[
    A \xrightarrow{\sigma}
    (\ul{X} \to B) \xrightarrow{\ul{X} \to \tau}
    (\ul X \to (\ul Y \to C)) \to
    ((\ul X \times \ul Y) \to C) \xrightarrow{\mu\to C}
    (\ul{X\times Y} \to C)
    \]
  We consider the composite of the first two morphisms.  
  We know that a play $t\in \sigma;(\X\to \tau)$ is given by $\s\vert_{A,X,Y,C}$, where $\s\vert_{A,X,B}\in\sigma$ and $\s\vert_{B,Y,C}\in\tau$.
  Write $\s_X$ for $\s\vert_{A,X,B}$ and $\s_Y$ for $\s\vert_{B,Y,C}$.

>>>>

\begin{definition}
  Given a set $X$ and a \Mellies morphism $\sigma \from 1 \to (X \to \bC)$, we say that $\sigma$ \emph{accepts} a sequence $u\in X^*$ if $\sigma;\xi_u\ne\bot$.
  We write $\Acc(\sigma)$ for the set of all sequences accepted by $\sigma$.
  \label{DefAcc}
\end{definition}

@@

\begin{proposition}
  Let $\sigma\from 1 \to (j(p) \to A)$, $\tau\from 1 \to (j(q) \to A)$ be two representatives of the same morphism $1 \to A$ in $\G/\oppcat\X$.  
  Then $(p,\Acc(\sigma))$ and $(q,\Acc(\tau)$ are equivalent.
  \label{PropAccEquivalent}
\end{proposition}
\begin{proof}
  Since the relation on pairs $(p,\U)$ is an equivalence relation, it suffices to assume that $\sigma$ and $\sigma'$ are related by the relation that generates the equivalence relation on \Mellies morphisms; i.e., that there is a morphism $f\from p \to q$ such that $\sigma = \tau;(j(f)\to \bC)$.

  Let $u\in j(p)^*$.  
  Then we have
  \begin{IEEEeqnarray*}{rCl}
    u\in\Acc(\sigma) & \Leftrightarrow & \sigma;\xi_u\ne\bot \\
    & \Leftrightarrow & \tau;(j(f)\to A);\xi_u\ne\bot \\
    & \Leftrightarrow & \tau\xi_{j(f)_*u}\ne\bot \\
    & \Leftrightarrow & j(f)_*u \in \Acc(\tau)\,.
  \end{IEEEeqnarray*}
  Therefore, $(p,\Acc(\sigma))$ and $(q,\Acc(\tau))$ are equivalent.
\end{proof}

@@

\section{Game Semantics and Probability}

We now specialize to the case that $\G$ is the category of arenas and single-threaded strategies that we developed in Chapters \ref{ChapGames} and \ref{ChapFullAbstraction}.
This will allow us to capture the close relationship between the sequences $u$ that we have been considering and the plays in a strategy.

\begin{definition}
  Let $X$ be a set and let $u\in X^*$ be a sequence.  
  Consider $X$ as an arena $\ul X$.  
  Then we write $qu$ for the play in $\ul X$ given by
  \[
    q\,u^{(1)}\,\cdots\,q\,u^{(|u|-1)}\,.
    \]
  Note that any $P$-position in $\ul X$ is of the form $qu$ for some sequence $u$.  
\end{definition}

\begin{definition}
  Let $A$ be an arena, let $V$ be a random variable taking values in a set $X$, and let $\sigma \from X \to A$ be a single-threaded strategy.  
  We may consider $X$ as a game $\ul X$.  
  Let $s$ be a legal play of $A$.  
  If $t\in\sigma$, we write $t/s$ if $t\vert_A = s$ and if the last move of $t$ is the last move of $s$ (this implies in particular that $t\vert_{\ul X}$ is a $P$-position in $\ul X$).
  Then we define
  \[
    \Acc_s(\sigma) = \{u\in X^*\suchthat \exists t\in\sigma\esuchthat t/s,\,t\vert_{\ul X} = qu\}\,.
    \]
  We define
  \[
    \bP_V(s\in\sigma) = \bP(V, \Acc_s(\sigma))\,.
    \]
\end{definition}

We would like use this definition to define $\bP(s\in\sigma)$ for $\sigma$ a morphism in $\G/\oppcat{(\Rv_{\Omega}^{FS})}$, but we first need to check that this is well-defined with respect to the equivalence relation on \Mellies morphisms.

\begin{proposition}
  Let
  \begin{mathpar}
    (V\from \Omega \to X, \sigma \from \ul X \to A)
    \and
    (W\from \Omega \to Y, \sigma \from \ul Y \to A)
  \end{mathpar}
  be representatives of the same morphism $A \to B$ in $\G/\oppcat{(\Rv_{\Omega}^{FS})}$, where $X$ and $Y$ are sets.  
  Let $s$ be a legal play of $A$.
  Then $(V,\Acc_s(\sigma))$ and $(W,\Acc_s(\sigma'))$ are equivalent as pairs in the sense of Definition \ref{DefEquivalenceOfPairs}.
\end{proposition}
\begin{proof}
  It suffices by induction to prove this in the case that the two representatives are related by the relation that generates the equivalence relation on morphisms; i.e., that there is a probability-preserving function $f\from X \to Y$ such that $\sigma=\sigma';(j(f) \to A)$.

  Let $u\in X^*$.  
  Then we have
  \begin{IEEEeqnarray*}{rCl}
    u \in \Acc_s(\sigma) & \Leftrightarrow & \exists t\in \sigma \esuchthat t/s,\,t\vert_{\ul X}=qu \\
    & \Leftrightarrow & \exists t \in \sigma';(j(f) \to A) \esuchthat t/s,\,t\vert_{\ul X}=qu \\
    & \Leftrightarrow & \exists t' \in \sigma' \esuchthat t/s,\,t\vert_{\ul Y} = q(f_*u) \\
    & \Leftrightarrow & f_*u \in \Acc_s(\sigma')\,.
  \end{IEEEeqnarray*}
  Therefore, $(V,\Acc_s(\sigma))$ and $(W,\Acc_s(\sigma'))$ are equivalent.
\end{proof}

It follows by Proposition \ref{PropProbabilityWellDefined} that $\bP_V(s\in\sigma)=\bP_W(s\in\sigma')$ for all $s$.  
Therefore, the following is well-defined.

\begin{definition}
  Let $\sigma \from A \to B$ be a morphism in $\G/\oppcat{(\Rv_{\Omega}^{FS})}$, where $\G$ is the category of arenas and single-threaded strategies, and suppose that $\sigma$ is given (after currying) by a morphism
  \[
    \tilde\sigma \from \ul X \to (A \to B)
    \]
  in $\G$, together with a random variable $V$ taking values in $X$.
  Then we define
  \[
    \bP(s\in\sigma) = \bP_V(s\in\tilde\sigma)\,.
    \]
\end{definition}

\begin{definition}
  Let $\sigma,\sigma'\from A \to B$ be morphisms in $\G/\oppcat{(\Rv_{\Omega}^{FS})}$, where $\G$ is the category of arenas and single-threaded strategies.  
  We say that $\sigma\approx_{\bP}\sigma'$ if for all legal plays $s$ of $A \to B$ we have
  \[
    \bP(s\in\sigma) = \bP(s\in\sigma')\,.
    \]
\end{definition}

We now relate this definition to \Mellies composition of strategies.

\begin{definition}
  Let $A,B,C$ be arenas and let $s$ be a play in $A\to C$.  
  We write
  \[
    \wit_B(s) = \{\s\in\Int(A,B,C)\suchthat \s\vert_{A,C}=s\}\,.
    \]
\end{definition}

\begin{proposition}
  Let $\sigma\from A \to B$, $\tau\from B \to C$ be morphisms in $\G/\oppcat{(\Rv_{\Omega}^{FS})}$.  
  Let $s$ be a legal play in $A \to C$.
  Then
  \[
    \bP(s\in\sigma;\tau) = \sum_{\s\in\wit_B(s)}\bP(\s\vert_{A,B}\in\sigma)\bP(\s\vert_{B,C}\in\tau)\,.
    \]
\end{proposition}
\begin{proof}
  Suppose that $\sigma$ and $\tau$ are given by (equivalence classes of) pairs
  \begin{mathpar}
    (V\from\Omega\to X, \tilde \sigma \from A \to (\ul X \to B))
    \and
    (W\from\Omega\to Y, \tilde \tau \from B \to (\ul Y \to C))\,,
  \end{mathpar}
  where $V$ and $W$ are random variables and $\tilde \sigma,\tilde\tau$ are strategies in $\G$.
  Then the composition $\sigma;\tau$ in $\G/\oppcat{(\Rv_{\Omega}^{FS})}$ is given by $V\tensor W\from\Omega \to X\times Y$, together with the \Mellies composition
  \[
    A \xrightarrow{\tilde\sigma}
    (\ul{X} \to B) \xrightarrow{\ul{X} \to \tilde\tau}
    (\ul X \to (\ul Y \to C)) \to
    ((\ul X \times \ul Y) \to C) \xrightarrow{\mu\to C}
    (\ul{X\times Y} \to C)\,,
    \]
  where $\mu$ is $\langle\ul{\pr_X},\ul{\pr_Y}\rangle$.

  First suppose that $s$ is a sequence in $A \to C$, and let $\s\in\wit_B(s)$.  
  Let $\t$ be a sequence in $\sigma\|(\ul X \to \tau)$ such that $\t\vert_{A,B,C}=\s$.  
  Then $\t\vert_{A,\ul X \to B}\in\sigma$ and $\t\vert_{B,\ul Y,C}\in\tau$.
\end{proof}

We want to prove that this is preserved by composition of strategies.  
First, we prove a lemma.

  Let $s\in\sigma\semicom_{\text{Mell}}\tau$.  
  Then there exist four sequence $s_X,s_Y,\s_X,\s_Y$ such that
  \begin{itemize}
    \item $s$ is an interleaving of the sequences $s_X$ and $s_Y$;
    \item $\s_X\in\sigma$ and $\s_Y\in\tau$;
    \item $s_X\vert_{A,X} = \s_X\vert_{A,X}$; and
    \item $s_Y\vert_{C,Y} = \s_Y\vert_{C,Y}$.
  \end{itemize}
  \label{LemMelliesDecomposition}
\end{lemma}
\begin{proof}
  Recall that $\sigma\semicom_{\text{Mell}}\tau$ is given by the following composite in $\G$, where $\mu$ is the composite $\langle\ul{\pr_X},\ul{\pr_Y}\rangle$.
  \[
    A \xrightarrow{\sigma}
    (\ul{X} \to B) \xrightarrow{\ul{X} \to \tau}
    (\ul X \to (\ul Y \to C)) \to
    ((\ul X \times \ul Y) \to C) \xrightarrow{\mu\to C}
    (\ul{X\times Y} \to C)
    \]
  We consider the composite of the first two morphisms.  
  We know that a play $t\in \sigma;(\ul X\to \tau)$ is given by $\s\vert_{A,X,Y,C}$, where $\s\vert_{A,X,B}\in\sigma$ and $\s\vert_{B,Y,C}\in\tau$.
  Write $\s_X$ for $\s\vert_{A,X,B}$ and $\s_Y$ for $\s\vert_{B,Y,C}$.

@@
